---
title: "Genetic map construction"
author: "MichelMoser"
date: "April 24, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='/media/mmoser/data1/P.EXSERTA/PEXV1.1.3/genetic_maps/markers/F7-K/')
```

## Clustering of GBS markers to linkage groups


```{r loading, include=FALSE, cache=FALSE}
library(ggplot2)
library(qtlTools)
library(reshape2)
library(dplyr)
library(ASMap)
library(ABHgenotypeR)
library(qtl)
```

------
         
         
#### **preliminary analysis:** 

Inspect missings in markers and individuals

#### **missing data per individual**

```{r prefiltering1, message=FALSE, warning=FALSE}
#missing data per marker
mrk <- read.csv2("LM3_F7_K.MARKER.genoSUMMARY.csv", header = T, sep = ",")

head(mrk)
mrks <- melt(mrk)
mrks <- mrks %>% filter(variable != "tot")

#check for exessive missings
mrk[which(mrk$NA. > 100),]

ggplot(data = mrks, aes(value, fill = variable)) +
  geom_density(alpha = 0.2)

```

#### **missing data per individual**

```{r prefiltering2, message=FALSE, warning=FALSE}
#missing data per individual
indv <- read.csv2("LM3_F7_K.INDIVIDUAL.genoSUMMARY.csv", header = T, sep = ",")
indvs <- melt(indv)
head(indv)
head(indvs)
indvs <- indvs %>% filter(variable != "tot")

#distribution of missings and ax and ex genotypes
ggplot(data = indvs, aes(value, variable, color = variable)) +
  geom_point(alpha = 0.2)
#ax vs ex
ggplot(data = indv, aes(AX, EX)) +
  geom_point(alpha = 0.2)
#missings
ggplot(data = indv, aes(NA.)) +
  geom_histogram(alpha = 0.2)

#remove those 4 individuals from the genetic map building

indv[which(indv$NA. > 1000),]

# 4 individuals removed and altered as LM3_F7_K.markers.clean.csv
```

# **Map building** 


```{r map initialising, message=FALSE, warning=FALSE}
f7_K <- read.cross(format = "csv", file = "LM3_F7_K.markers.clean.csv", F.gen = 7, genotypes = c("AX", "HET", "EX"))

f7_K <- convert2riself(f7_K)

#save duplicates and corresponding markers in bins to reconstruct them later

duplicatesf7_K <- findDupMarkers(f7_K)

f7_dups <- do.call(rbind, lapply(seq_along(duplicatesf7_K), function(i){
  data.frame(bin_name=names(duplicatesf7_K)[i], duplicatesf7_K[[i]])}))
head(f7_dups)
write.table(f7_dups,file = "LM3_F7_K.marker.dupbins1.txt", quote = FALSE, row.names = FALSE)

#remove duplicate markers
f7_K <- pullCross(f7_K, type = "co.located")



```
# Generate inital map

```{r map_construction, message=FALSE, warning=FALSE}

#set bychr = FALSE to allow complete reconstruction of map 
map1 <- mstmap(f7_K, pop.tpye = "ARIL", bychr = F, dist.fun = "kosambi", trace = TRUE, detectBadData = T, p.value = 1e-09, , mvest.bc = TRUE, return.imputed = T)

#sort markers in each LG
map1 <- mstmap(map1, pop.tpye = "ARIL", bychr = T, dist.fun = "kosambi", trace = TRUE, detectBadData = T, p.value = 1e-09, , mvest.bc = TRUE, return.imputed = T)

#inspect results
summary(map1)

geno.image(map1)


head(geno.table(map1))
#geno.image(map1, col = c("white", "darkorange", "lightblue"))
heatMap(map1, lmax = 25)
nmar(map1)
summary(map1)



```

## Merging of unlinked markers

Expected 7 LGs in Petunia, therefore merge small LGs into a single one and try to push markers onto the 7 large LGs

```{r merge}
#put all small LGs in to one
nmar(map1)

row.names(as.data.frame(unlist(nmar(map1)[nmar(map1) < 20])))

map2 <- mergeCross(map1, merge = list(
                       "UNKNOWN" =row.names(as.data.frame(unlist(nmar(map1)[nmar(map1) < 20])))
))
heatMap(map2)
#try to assign them to chromosomes
#map2 <- pushCross(map2, type = "unlinked", unlinked.chr = "UNKNOWN" )

#map3 <- mstmap(map2, bychr = TRUE, trace = TRUE, anchor = TRUE, p.value= 2)
#summary(map3)
nmar(map1)

# summary(map2)
#no improvement

names(map2$geno)[8] <- "L.8"

profileGen(map2, bychr = FALSE, stat.type = c("xo", "dxo", "miss"), id = "Genotype", xo.lambda = 25, layout = c(1, 3), lty = 2)

write.cross(format = "csv", map2, "F7_Kmap2.asmap")
?profileGen
```

# Conversion to ABHgenotyper

```{bash}
python ~/Tools/repositories/GBS2map/Rqtl_2_abh.py -i F7_Kmap2.asmap.csv -o F7_Kmap2.abhIN.csv -q2a
```

# ABH genotyper 




```{r}
#test 
library(ABHgenotypeR)
abh <- readABHgenotypes("F7_Kmap2.abhIN.csv", nameA = "A", nameB = "B")

plotGenos(abh)
# NA's are mistakenly labelled as heterozygotes when heterozygotes are absent. Bug in abhgenotyper code

pes <- imputeByFlanks(abh)

?correctStretches
pes <- correctStretches(pes, maxHapLength = 2)
pes <- imputeByFlanks(pes)

reportGenos(pes)
plotGenos(pes, chromToPlot = 1:7)
# NOTE: 

# NA's are mistakenly labelled as heterozygotes when heterozygotes are absent. Bug in abhgenotyper code

writeABHgenotypes(pes, "F7_Kmap3.abhOUT.csv")
```

### convert back to Rqtl format

```{bash}
python ~/Tools/repositories/GBS2map/Rqtl_2_abh.py -i F7_Kmap3.abhOUT.csv -o F7_Kmap3.asmapIN.csv -a2q


```

### rebuild map from imputed and corrected markers in ASMap


```{r}

f7_K2 <- read.cross(format = "csv", file ="F7_Kmap3.asmapIN.csv", F.gen = 7, genotypes = c("A", "HET", "B"))
f7_K2 <- convert2riself(f7_K2)
f7_K2 <- pullCross(f7_K2, type = "co.located")
f7_K2

#set bychr = FALSE to allow complete reconstruction of map 
map4 <- mstmap(f7_K2, pop.tpye = "ARIL", bychr = F, dist.fun = "kosambi", trace = TRUE, detectBadData = T, p.value = 1e-09, , mvest.bc = TRUE, return.imputed = T)

```

# Final map

```{r}


summary(map4)
heatMap(map4)
plot.map(map4)

profileGen(map4, bychr = FALSE, stat.type = c("xo", "dxo", "miss"), id = "Genotype", xo.lambda = 25, layout = c(1, 3), lty = 2)

profileMark(map4, stat.type = c("seg.dist", "dxo", "erf", "lod"), id = "Genotype", layout = c(1, 4), type = "l")

```

